General purpose gene annotation using deep learning

### 2020-04-07 paper/main.tex
lots of reordering

\begin{table}[!t]
\processtable{Data group statistics for our training data\label{Tab:statistics_training}} {
\begin{tabular}{@{}lll@{}}
\toprule & Animals & Plants\\
\midrule
Average genome size in Gbp& 2.489 (+- 2.073) & 0.914 (+- 0.934) \\
Average gene length & 25,672 (+- 16,605) & 3,509 (+- 906)\\
Geenuff error rate & 0.253 (+- 0.138) & 0.134 (+- 0.072) \\
Fraction of class Intergenic & 0.752 (+- 0.05) & 0.808 (+- 0.106) \\
Fraction of class UTR & 0.013 (+- 0.011) & 0.033 (+- 0.023) \\
Fraction of class CDS & 0.028 (+- 0.028) & 0.077 (+- 0.052) \\
Fraction of class Intron  & 0.207 (+- 0.023) & 0.083 (+- 0.037) \\
\botrule
\end{tabular}}{The description of Table \ref{Tab:statistics_all} applies here as well.}
\end{table}

During data generation, we only queried for the longest transcript of each gene and did not include FASTA sequences that had no functional annotation (explanation why this is a sign of error?). As Geenuff checks the genomic annotations for potential errors and is able to mark those areas, we were able to effectively blank out those bases during training. The vast majority of bases masked lie in the intergenic region as the most prevelant error is a missing UTR and Geenuff marks a potentially large intergenic region for it. Table \ref{Tab:statistics_all} and \ref{Tab:statistics_training} show statistics about the masking. For the training, we divided each continous genomic sequence into 20.000 bases long subsequences, which are then used as input or label for the neural network. We appended zero padding if the subsequences are not long enough. If a subsequence is fully marked as erroneous, we do not include it in the training or evaluation datasets.

@@

\begin{table}[!t]
\processtable{Data arrays generated and used by Helixer\label{Tab:01}} {
\begin{tabular}{@{}ll@{}}
\toprule Name & Information \\
\midrule
Input & Genomic sequence in a one hot encoding\\
	  & for (A, T, C, G), otherwise different, see suplementary .. ()\\
Output & Labels in a 4-dimensional one hot encoding \\
Sample weights & One of $\{0,1\}$; whether there is an error at a base\\
RNASeq score & A score in the interval $...$ to rank the coherence of the \\
		     & annotation with RNASeq information \\
\botrule
\end{tabular}}{}
\end{table}

text\vspace*{1pt}

\begin{itemize}
\item for bulleted list, use itemize
\item for bulleted list, use itemize
\item for bulleted list, use itemize\vspace*{1pt}
\end{itemize}

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text

\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.

@@

%\begin{figure}[!tpb]%figure2
%%\centerline{\includegraphics{fig02.eps}}
%\caption{Caption, caption.}\label{fig:02}
%\end{figure}

@@

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
\citealp{Boffelli03} might want to know about  text text text text

@@

We propose a framework to develop a gene calling model that is both more general in the range genomes it can annotate and also outperforms traditional methods on a base pair wise evaluation. The overall results of Helixer, Augustus and a baseline build with dilated convolutional layers are shown in table \ref{Tab:03}.

\section{Discussion}

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text.

Table~\ref{Tab:01} shows that Text Text Text Text Text  Text Text
Text Text Text Text. Figure~2\vphantom{\ref{fig:02}} shows that
the above method Text Text. Text Text Text  Text Text Text Text
Text Text. Figure~2\vphantom{\ref{fig:02}} shows that the above
method Text Text. Text Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method Text
Text.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%     please remove the " % " symbol from \centerline{\includegraphics{fig01.eps}}
%     as it may ignore the figures.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusion}

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text\vspace*{-10pt}

@@

Text Text Text Text Text Text  Text Text.  \citealp{Boffelli03} might want to know about  text
text text text\vspace*{-12pt}

@@

\end{thebibliography}
\end{document}

>>>>

During data generation, we only queried for the longest transcript of each gene and did not include FASTA sequences that had no functional annotation (explanation why this is a sign of error?). As Geenuff checks the genomic annotations for potential errors and is able to mark those areas, we were able to effectively blank out those bases during training. The vast majority of bases masked lie in the intergenic region as the most prevelant error is a missing UTR and Geenuff marks a potentially large intergenic region for it. Table \ref{Tab:statistics_all} shows statistics about the masking. The corresponding statistics for the selected training genomes can be found in Table \ref{Tab:statistics_training}. For the training, we divided each continous genomic sequence into 20.000 bases long subsequences, which are then used as input or label for the neural network. We appended zero padding if the subsequences are not long enough. If a subsequence is fully marked as erroneous, we do not include it in the training or evaluation datasets.

@@

\begin{table}[!t]
\processtable{Data arrays generated and used by Helixer\label{Tab:01}} {
\begin{tabular}{@{}ll@{}}
\toprule Name & Information \\
\midrule
Input & Genomic sequence in a one hot encoding\\
	  & for (A, T, C, G), otherwise different, see suplementary .. ()\\
Output & Labels in a 4-dimensional one hot encoding \\
Sample weights & One of $\{0,1\}$; whether there is an error at a base\\
\botrule
\end{tabular}}{}
\end{table}

@@

@@

@@

We evaluate the performance of our final model base on ... The overall results are shown in table \ref{Tab:03}.

\section{Conclusion}
We propose a gene calling model that is able to perform cross-species genomic annotiation for a wide variety of large eucaryotic genomes. Our model does this while outperforming AUGUSTUS on a base pair wise level and withstanding RNASeq data based scrutiny.

@@

We are very grateful for the support of...

@@

\end{thebibliography}

\newpage

\section{Supplementary Material}
\subsection{Dilated CNN Baseline Network Search}

\subsection{Input Data Encoding}

\subsection{Hyperparameters}

\subsection{Training Data statistics}
\begin{table}[!t]
\processtable{Data group statistics for our training data\label{Tab:statistics_training}} {
\begin{tabular}{@{}lll@{}}
\toprule & Animals & Plants\\
\midrule
Average genome size in Gbp& 2.489 (+- 2.073) & 0.914 (+- 0.934) \\
Average gene length & 25,672 (+- 16,605) & 3,509 (+- 906)\\
Geenuff error rate & 0.253 (+- 0.138) & 0.134 (+- 0.072) \\
Fraction of class Intergenic & 0.752 (+- 0.05) & 0.808 (+- 0.106) \\
Fraction of class UTR & 0.013 (+- 0.011) & 0.033 (+- 0.023) \\
Fraction of class CDS & 0.028 (+- 0.028) & 0.077 (+- 0.052) \\
Fraction of class Intron  & 0.207 (+- 0.023) & 0.083 (+- 0.037) \\
\botrule
\end{tabular}}{The description of Table \ref{Tab:statistics_all} applies here as well.}
\end{table}

\end{document}
